---
title: "Proportion of Brazil as the affiliation of documents in SciELO Brazil"
output:
  html_document:
    df_print: paged
---

Let's load from SciELO Analytics
the CSV of documents affiliations in SciELO Brazil:

```{r}
# We shouldn't interpret Namibia (NA) as "not available"
doc_aff <- read.csv("tabs_bra/documents_affiliations.csv", na.strings = c())
doc_aff
```

R already simplifies the column names in some sense,
replacing the whitespaces and special characters by a dot.
We can see the column names with:

```{r}
names(doc_aff)
```

Categorical fields are known as *factors*.

```{r}
class(doc_aff$document.type)
class(doc_aff$document.affiliation.country.ISO.3166)
```

The *levels* of a factor are the values one *factor* vector can have.

```{r}
levels(doc_aff$document.type)
levels(doc_aff$document.affiliation.country.ISO.3166)
```

Most entries are research articles,
we'll work only with this document type:

```{r}
summary(doc_aff$document.type)
par(mar = c(3, 9, 2, 2) + .1)
barplot(summary(doc_aff$document.type), horiz = TRUE, las = 1)
articles <- doc_aff[doc_aff$document.type == "research-article",]
articles
```

Most affiliation entries are from Brazil
(that's somewhat expected for a Brazilian collection).

```{r}
barplot(summary(articles$document.affiliation.country.ISO.3166,
                maxsum = 10))
```

Let's build a dataset with just four columns:

* One regarding the document publication year;
* One regarding to the PID, a way to identify an article;
* One logical, `TRUE` if an article have a Brazilian affiliation;
* One logical, `TRUE` if an article have a non-Brazilian affiliation.

We should remove the empty country entries,
since they might belong to any country (Brazil or other).
Using two columns should be cleaner to understand
than merging the Brazilian/non-Brazilian affiliation as a single column.

```{r}
dataset <- data.frame(
  articles$document.publishing.year,
  articles$document.publishing.ID..PID.SciELO.,
  articles$document.affiliation.country.ISO.3166 == "BR",
  grepl("[^B].|.[^R]", articles$document.affiliation.country.ISO.3166)
)
names(dataset) <- c("year", "pid", "br", "not_br")
dataset <- dataset[dataset$br | dataset$not_br,]
dataset
```

As all entries are either `br` or `not_br`,
we just need to calculate the mean of `br` for each PID.
We'll use `dplyr` to group that result by the PID.

```{r}
library(dplyr)
proportions <- dataset %>% group_by(pid) %>% summarize(mean(br), max(year))
proportions <- proportions[c(2, 3)]
names(proportions) <- c("prop", "year")
proportions
```

Let's see the evolution of the mean of these proportions:

```{r}
mprops <- proportions %>% group_by(year) %>% summarize(mean(prop))
mprops
plot(
  mprops,
  type = "l",
  cex.main = 1,
  main = paste("Mean proportion of BR affiliation",
               "in research articles (SciELO Brazil)",
               sep = " ")
)
```

Is that significantly decreasing?

```{r}
regr <- lm(mean.prop. ~ year, data.frame(mprops))
summary(regr)
confint(regr, level = .95)
```

Yes! The slope (`year`, last row of `confint` result)
is negative for the entire $95\%$ confidence interval.
